### 2019-08-21 Modeling.Rmd

---
title: "Single Cell Xenopus TMT Analysis: Modeling"
author: "Kyle Kovary"
date: "5/15/2019"
output: html_document
---

## Complex Stoichiometry
Could be that when drawing from a correlated distribution, the variance between cells will be the same, but the variance within a cell will be much less... High correlation increases variance between cells but reduces variance within a cell for a given pathway. Need to test that high variance with high covariance matters for a single cell, could be that correlation is all that matters, which would invalidate the reasoning that for certain pathways there is a tradeoff between variance and correlation.
```{r}
correlatedValue = function(x, r){
  r2 = r**2
  ve = 1-r2
  SD = sqrt(ve)
  e  = rnorm(length(x), mean=0, sd=SD)
  y  = r*x + e
  return(y)
}

#set.seed(5)
n = 1000
x = rnorm(n)
y = correlatedValue(x=x, r=.6)

# multiply the variable by the SD you want and then add the mean you want

corMat <- function(nRow = 1000, nCol = 10, R = 0.5, mean = 1000, sd = 30){
  mat <- matrix(data = rnorm(nRow), nrow = nRow)

  for(i in 1:(nCol) - 1){
    mat <- cbind(mat, correlatedValue(x = mat[,i], r = R))
  }

  return((mat*sd) + mean)
}

```

```{r}
library(mvtnorm)

fun = function(cor_matrix, list_distributions, L)
{
    n = length(list_distributions)
    # Correlated Gaussian variables
    Gauss = rmvnorm(n=L, mean = rep(0,n), sig=cor_matrix)
    # convert them to uniform distribution.
    Unif = pnorm(Gauss) 
    # Convert them to whatever I want
    vars = sapply(1:n, FUN = function(i) list_distributions[[i]](Unif[,i]))
    return(vars)
}

# Create a matrix of values with a given R, mean, and standard deviation
corProts <- function(nProt, nCell, R, mean, sd) {

  cor_matrix =  matrix(data = R,
                       ncol = nProt,
                       nrow = nProt)
  for (i in 1:nProt) {
    cor_matrix[i, i] = 1
  }

  list_distributions <- list()
  for (i in 1:nProt) {
    list_distributions <-
      c(list_distributions, function(nCell)
        round(qnorm(nCell, mean, sd)))
  }

  mat = fun(cor_matrix, list_distributions, nCell)
  return(mat)
}

# Plot of total variance of 10 subunit complex

# Plot of max number of complete complexes
cors <- seq(0,1,0.02)
vars <- rev(seq(0,1,0.02))

corVarMat <- matrix(data = NA, ncol = length(cors), nrow = length(vars))
colnames(corVarMat) <- cors
rownames(corVarMat) <- vars

# Percent of potentially complete complexes
for(i in 1:length(cors)){
  for(j in 1:length(vars)){
    perc <- corProts(nProt = 10, nCell = 10000, R = cors[i], mean = 1000, sd = 1000*vars[j])
    perc <- mean(apply(perc, 1, function(x) min(abs(x)) / max(abs(x))), na.rm = T)
    corVarMat[j,i] <- perc
  }
}
pheatmap(corVarMat, cluster_cols = F, cluster_rows = F)
rotate <- function(x) t(apply(x, 2, rev))
pheatmap(rotate(corVarMat), cluster_cols = F, cluster_rows = F)
plot(diag(rotate(corVarMat)), rotate(corVarMat)[])

x <- tibble(index = 1:51,
            diagnal = diag(corVarMat), 
            zeroCor = corVarMat[,1],
            fullCor = corVarMat[1,])
x %>% gather(key = slice, value = percent, 2:4) %>% 
  ggplot(., aes(x = index, y = percent, colour = slice)) + 
  geom_point() + geom_vline(xintercept = 40, linetype = 'dashed')

x <- as_tibble(cbind(vars, corVarMat))
x <- x %>% gather(key = R, value = percent, 2:ncol(x)) %>% mutate(R = as.numeric(R)) %>%
  rename_(CV = vars)
ggplot(x, aes(x = CV, y = percent, colour = R, group = as.factor(R))) + geom_path() + scale_x_reverse() + theme_bw()

ggplot(x, aes(x = CV, y = percent, colour = as.factor(R))) + geom_path() + scale_x_reverse(limits = c(0.2,0)) + theme_bw()

# Within cell variance
for(i in 1:length(cors)){
  for(j in 1:length(vars)){
    stdev <- corProts(nProt = 10, nCell = 10000, R = cors[i], mean = 1000, sd = 1000*vars[j])
    stdev <- median(apply(stdev, 1, function(x) sd(x, na.rm = T) / mean(x, na.rm = T)), na.rm = T)
    corVarMat[j,i] <- stdev
  }
}

```