### 2020-02-25 Analysis.Rmd
updated

---
title: "Single Cell Xenopus TMT Analysis"
author: "Kyle Kovary"
date: "5/15/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(RColorBrewer)
library(pheatmap)
library(GGally)
library(VIM)
library(org.Xl.eg.db)
library(GO.db)
library(ggrepel)
library(plotly)

# Import Data
files = list.files('TMTdata/', pattern = 'TargetPeptideSpectrumMatch', full.names = T)

for(i in seq_along(files)){
  if(i == 1 & exists('pepRaw')){
    rm(pepRaw)
  } else if(!exists('pepRaw')){
    pepRaw = read_tsv(file = files[i]) %>% mutate(file = files[i], massID = 1:nrow(.))
  } else{
    temp = read_tsv(file = files[i]) %>% mutate(file = files[i], massID = 1:nrow(.))
    colnames(temp) = colnames(pepRaw)
    pepRaw = rbind(pepRaw, temp)
    rm(temp)
  }
}
pepRaw <- pepRaw %>% dplyr::select(file, massID, everything()) %>% filter(!is.na(`126`), 
                                                                  `Quan Usage` == 'Use',
                                                                  Confidence != 'Low')

```

## Data Organization

```{r }
pepRaw = pepRaw %>% 
  mutate(run = c('ctrl','a','b','c','d','e')[match(pepRaw$file, files)]) %>% 
  group_by(file, `Annotated Sequence`) %>% 
  mutate(pepID = letters[1:n()]) %>% ungroup() %>%
  dplyr::select(file, run, pepID, everything())

pepTidy = pepRaw %>% 
  gather(key = label, value = intensity, '126':'131') %>% 
  dplyr::select(run, label, intensity, everything(), -file) %>% 
  mutate(condition = ifelse(run == 'ctrl', 'ctrl',c('ctrl',0,20,40,60,80)[match(label,c(126:131))]),
         intensity = as.numeric(intensity)) %>% unite(peptide, c(`Annotated Sequence`,'pepID')) %>%
  dplyr::rename(Uniprot = `Master Protein Accessions`) %>% 
  unite(cell, c('run','label','condition'), remove = F) %>%
  dplyr::select(cell, run, label, condition, peptide, Uniprot, intensity)
```

## Variance and Co-variance Between Cells Before Normalization

```{r echo=TRUE, fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
# ggplot(pepTidy, aes(log2(intensity), colour = cell)) + 
#   geom_density() + theme_bw() + xlim(-20,20)
```

The peptide intensities appear to follow a log-normal distribution.

```{r echo=TRUE, fig.height=10, fig.width=10,message=FALSE, warning=FALSE}
pairs = pepTidy %>% filter(condition == 'ctrl') %>% 
  dplyr::select(cell, peptide, intensity) %>% spread(cell, intensity)

# ggpairs(log2(pairs[,2:ncol(pairs)]), 
#         lower = list(continuous = wrap("points", alpha = 0.1, size = 0.1))) + theme_bw()
```

```{r echo=TRUE, fig.height=10, fig.width=10,message=FALSE, warning=FALSE}
meanSD = pepTidy %>% filter(condition != 'ctrl') %>% 
  group_by(peptide) %>% 
  summarise(mean = mean(intensity, na.rm = T), 
            sd = sd(intensity, na.rm = T) / mean)

# ggplot(meanSD, aes(x = log10(mean), y = sd)) + geom_point() + geom_smooth(method = lm, formula = y ~ poly(x, 2))
```

## Variance and Co-variance Between Cells After Cell Normalization

```{r echo=TRUE, fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
pepTidy = pepTidy %>% group_by(cell) %>% mutate(intensity = intensity / median(intensity, na.rm = T))

# ggplot(pepTidy, aes(log2(intensity), colour = cell)) + geom_density() + theme_bw() + xlim(-8,8)
```

```{r echo=TRUE, fig.height=10, fig.width=10,message=FALSE, warning=FALSE}
pairs = pepTidy %>% filter(condition == 'ctrl') %>% 
  dplyr::select(cell, peptide, intensity) %>% spread(cell, intensity)

# ggpairs(log2(pairs[,2:ncol(pairs)]), lower = list(continuous = wrap("points", alpha = 0.1, size = 0.1))) + theme_bw()
```

## Variance and Co-variance Between Cells After Peptide Batch Normalization

```{r echo=TRUE, fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
pepTidy = pepTidy %>% group_by(run, peptide) %>% 
  mutate(intensity = intensity / median(intensity, na.rm = T)) %>% ungroup()

# ggplot(pepTidy, aes(log2(intensity), colour = cell)) + geom_density() + theme_bw() + xlim(-8,8)
```

```{r echo=TRUE, fig.height=10, fig.width=10,message=FALSE, warning=FALSE}
pairs = pepTidy %>% dplyr::filter(condition == 'ctrl') %>% 
  dplyr::select(cell, peptide, intensity) %>% spread(cell, intensity)

# ggpairs(log2(pairs[,2:ncol(pairs)]), lower = list(continuous = wrap("points", alpha = 0.1, size = 0.1))) + theme_bw()
```

There are

## Variance and Co-variance Between Cells After Removing Peptides Outside of 2*sd

```{r echo=TRUE, fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
pepTidy %>% dplyr::filter(abs(log2(intensity)) < 2*sd(log2(pepTidy$intensity), na.rm = T))

ggplot(pepTidy, aes(log2(intensity), colour = cell)) + geom_density() + theme_bw() + xlim(-4,4)
```

The standard deviation of all peptides across eggs was calculated, and every peptide that had a value outside of `r 2*sd(log2(pepTidy$intensity), na.rm = T)`

## Group into proteins based on median

```{r echo=TRUE, fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
protTidy = pepTidy %>% filter(abs(log2(intensity)) < 2*sd(log2(pepTidy$intensity), na.rm = T)) %>% group_by(cell, run, label, condition, Uniprot) %>% 
  summarise(abundance = median(intensity, na.rm = T)) %>% ungroup()

pairs = protTidy %>% dplyr::filter(condition == 'ctrl') %>% 
  dplyr::select(cell, Uniprot, abundance) %>% 
  spread(cell, abundance)

ggpairs(log2(pairs[,2:ncol(pairs)]), lower = list(continuous = wrap("points", alpha = 0.1, size = 0.1))) + theme_bw()

pairs = protTidy %>% dplyr::select(cell, Uniprot, abundance) %>% spread(cell, abundance)

protImpute = kNN(data = pairs, k = 1) %>% as_tibble() %>% dplyr::select(Uniprot:e_131_80)

cor = cor(t(log2(protImpute[c(1:ncol(protImpute))[-c(1,grep('ctrl',colnames(protImpute)))]])), use = 'pairwise.complete.obs')
colnames(cor) = protImpute$Uniprot
rownames(cor) = protImpute$Uniprot

mypal = rev(colorRampPalette(brewer.pal(11, 'RdBu'))(201))
k = 2
map = pheatmap(cor, show_colnames = F, show_rownames = F, color = mypal, breaks = seq(-1,1, by = 0.01), cutree_cols = k, cutree_rows = k, legend = F)
#map

tibble(Uniprot = protImpute$Uniprot, Cluster = cutree(map$tree_row, k = k)) %>% write_csv('/Users/kylekovary/Downloads/Map.csv')

remove = which(is.na(pairs), arr.ind = T)[,1] %>% unique()
pca = prcomp(protImpute[,2:ncol(protImpute)])

pcaDF = tibble(cell = rownames(pca$rotation), PC1 = pca$rotation[,1], PC2 = pca$rotation[,2])
pcaDF = pcaDF %>% separate(cell, into = c('run','label','condition'), remove = F) %>% print()
ggplot(pcaDF, aes(x = PC1, y = PC2, colour = condition)) + geom_point(size = 5)

```

## Map GO Terms to Proteins

```{r eval=F}
columns(org.Xl.eg.db)
#mapIds(org.Xl.eg.db, keys = 'A0JPG5', keytype = 'UNIPROT', column = 'GOALL', multiVals = 'list')

keys <- keys(org.Xl.eg.db, keytype = 'UNIPROT')

GO <- protImpute %>% dplyr::select(Uniprot) %>% filter(Uniprot %in% keys) %>% 
  mutate(GOterms = mapIds(org.Xl.eg.db, 
                     keys = Uniprot, 
                     keytype = 'UNIPROT', 
                     column = 'GOALL', 
                     multiVals = 'list'))

allGO <- unnest(GO, cols = GOterms) %>% nest(Uniprot = Uniprot) %>%
  rowwise() %>% mutate(n = length(unlist(Uniprot))) %>% ungroup() %>%
  mutate(GO = mapIds(GO.db, 
                     keys = GOterms, 
                     keytype = 'GOID', 
                     column = 'TERM') %>% as.character(),
         ontology = mapIds(GO.db, 
                     keys = GOterms, 
                     keytype = 'GOID', 
                     column = 'ONTOLOGY') %>% as.character()) %>%
  dplyr::select(GOterms, GO, ontology, everything())

# allGO <- tibble(GOterms = unlist(GO$GO) %>% unique())
# allGO <- allGO %>% group_by(GOterms) %>% mutate(Uniprot = GO[grep(GOterms, GO$GO),1] %>% as.list()) %>% ungroup() %>%
#   mutate(
#          GO = mapIds(GO.db, 
#                      keys = GOterms, 
#                      keytype = 'GOID', 
#                      column = 'TERM') %>% as.character(),
#          ontology = mapIds(GO.db, 
#                      keys = GOterms, 
#                      keytype = 'GOID', 
#                      column = 'ONTOLOGY') %>% as.character()) %>%
#   dplyr::select(GOterms, GO, ontology, everything())

corVector <- function(uniprotList, corMat){
  targets = uniprotList %>% unlist()
  temp = corMat[rownames(corMat) %in% targets,colnames(corMat) %in% targets]
  temp[lower.tri(temp, diag = F)]
}

selectCV <- function(x){
  protVar %>% filter(Uniprot %in% unlist(x)) %>% 
    dplyr::select(cv) %>% as_tibble() %>% as.list()
}

allGO <- allGO %>% group_by(GOterms) %>% 
  dplyr::mutate(cors = corVector(Uniprot, cor) %>% as_tibble() %>% as.list(),
                cv = selectCV(Uniprot)) %>% ungroup() %>%
  rowwise() %>% mutate(medCors = mean(cors, na.rm = TRUE),
                       medCV = mean(cv, na.rm = TRUE)) %>%
  ungroup() %>%
filter(length(Uniprot) > 3,
       !is.na(medCors),
       !is.na(medCV))

#saveRDS(GO, 'analysisData/GO.RDS')
#saveRDS(allGO, 'analysisData/allGO.RDS')
```

## GO Term Analysis
```{r echo=TRUE, message=FALSE, warning=FALSE}
#GO <- readRDS('analysisData/GO.RDS')
#allGO <- readRDS('analysisData/allGO.RDS')

x <- allGO %>% rowwise %>% mutate(size = nrow(Uniprot)) %>% dplyr::select(GOterms, size, cors) %>% unnest() %>% group_by(GOterms) %>% 
  summarise(median = median(cors, na.rm = T),
            sd = sd(cors, na.rm = T),
            size = mean(size)) %>% 
  mutate(GOterms = fct_reorder(GOterms, median))

ggplot(filter(x, size >= 2), aes(x = GOterms, y = median)) + geom_point(alpha = 0.25) +
  geom_hline(yintercept = 0.02154303, colour = 'red', linetype = 'dashed') + 
  theme_bw() + ylab('Mean Correlation') + ylim(-1,1)

```

## Plot highest correlated GO Terms with GO Term Names
```{r}
topGOnames <- x %>% filter(size > 2) %>% .[order(.$median, decreasing = T),] %>% 
  .[1:50,'GOterms'] %>% unlist()

topGO <- allGO %>% filter(GOterms %in% topGOnames) %>% 
  dplyr::select(GOterms, GO, cors) %>% unnest() %>%
  group_by(GO) %>% mutate(median = median(cors)) %>% ungroup() %>%
  mutate(GO = fct_reorder(GO, median))

ggplot(topGO, aes(x = GO, y = cors)) + geom_boxplot() + geom_jitter(alpha = 0.5, size = 0.5) + coord_flip() + scale_x_discrete(labels = function(x) str_wrap(x, width = 50)) + theme_bw() + ylim(-1,1)

mitoProts <- allGO %>% filter(GOterms == 'GO:0051321') %>% dplyr::select(Uniprot) %>% unnest() %>% unlist()
mitoDF <- protTidy %>% filter(condition != 'ctrl', Uniprot %in% mitoProts) %>% group_by(condition, Uniprot) %>% summarise(mean = mean(abundance), sd = sd(abundance) / (n()-1))

ggplot(mitoDF, aes(x = as.numeric(condition), y = mean, colour = Uniprot)) + geom_point() + geom_path() + geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0, alpha = 0.5) + theme_bw()

```

## Variance of GO Terms
```{r}

protVar <- protImpute %>% gather(key = 'cell', value = 'abundance', 2:ncol(protImpute)) %>%
  separate(cell, into = c('run','label','condition'), remove = F) %>% 
  filter(condition != 'ctrl') %>% group_by(Uniprot) %>%
  summarise(cv = sd(abundance)/mean(abundance))

ggplot(protVar, aes(x = cv)) + geom_histogram(bins = 50) + theme_bw()

meanCV = allGO %>% dplyr::select(GOterms, cv) %>% unnest() %>% group_by(GOterms) %>% 
  summarise(mean = mean(cv), size = n()) %>% mutate(GOterms = fct_reorder(GOterms, mean))

ggplot(meanCV, aes(x = GOterms, y = mean, size = log2(size), alpha = 0.2)) + 
  geom_point() + theme_bw()

topGOnames <- meanCV %>% filter(size > 2) %>% .[order(.$mean, decreasing = F),] %>% 
  .[1:40,'GOterms'] %>% unlist()

topGO <- unnest(dplyr::select(filter(allGO, GOterms %in% topGOnames), GOterms, cv)) %>%
  mutate(GO = mapIds(GO.db, keys = GOterms, keytype = 'GOID', column = 'TERM')) %>%
  group_by(GO) %>% mutate(median = median(cv)) %>% ungroup() %>%
  mutate(GO = fct_reorder(GO, median))

ggplot(topGO, aes(x = GO, y = cv)) + geom_boxplot(outlier.size = 1) + geom_point() + coord_flip() + scale_x_discrete(labels = function(x) str_wrap(x, width = 50)) + theme_bw() + ylim(0,0.3)
```

```{r message=FALSE, warning=FALSE}
allGO[grep('ribosom',allGO$GO),]
glyc <- filter(allGO, GOterms == 'GO:0005840')$Uniprot %>% unlist()

glycMat <- cor[rownames(cor) %in% glyc,colnames(cor) %in% glyc]
rownames(glycMat) <- mapIds(org.Xl.eg.db, keys = rownames(glycMat), keytype = 'UNIPROT', column = 'SYMBOL')
pheatmap(glycMat, show_colnames = T, show_rownames = T, color = mypal, breaks = seq(-1,1, by = 0.01),
         cutree_cols = 3, cutree_rows = 3)

glyc <- protImpute %>% filter(Uniprot %in% glyc) %>% 
  mutate(symbol = mapIds(org.Xl.eg.db, keys = Uniprot, keytype = 'UNIPROT', column = 'SYMBOL')) %>%
  dplyr::select(Uniprot, symbol, everything())

glyc <- glyc %>% gather(key = sample, value = abundance, 3:ncol(glyc)) %>%
  dplyr::select(sample, symbol, abundance) %>% spread(key = symbol, value = abundance)

#ggpairs(log2(glyc[,2:ncol(glyc),]))
```

## Cor vs CV
```{r}
test = allGO %>% filter(length(Uniprot) > 3, !is.na(medCors), !is.na(medCV))

ggplot(test, aes(x = medCors, y = medCV)) + geom_point(alpha = 0.5) + theme_bw()

ggplot(test, aes(x = medCors, y = medCV)) + 
  geom_point(alpha = 0.5) + 
  #stat_smooth(method="lm",formula=  y ~ I(1/x)) +
  geom_text_repel(data = filter(test, medCors > 0.15), aes(label = GO), colour = 'black', size = 3) + 
  theme_bw()

test2 = tibble(Cor = test$medCors, CV = test$medCV, mult = test$medCors * test$medCV) %>%
  gather(key = 'Key', value = 'Value')
ggplot(test2, aes(x = Value, colour = Key)) + geom_density() + theme_bw()
ggplot(test2, aes(x = Key, y = Value)) + geom_boxplot() + theme_bw()
# Variance vs sum of covariance
# Cor vs Var relationship for every protein pair... error propagation
```

## All combinations of 5 proteins -> total error vs correlation
```{r}
combMat <- combn(1:length(colnames(cor)),2)
corVect <- tibble(protA = rep(rownames(cor), times = length(colnames(cor))), 
                  protB = rep(colnames(cor), each = length(rownames(cor))),
                  keep = as.vector(lower.tri(cor)),
                  cor = as.vector(cor)) %>% filter(keep) %>% dplyr::select(-keep)

corVect <- corVect %>% dplyr::rename(Uniprot = protA) %>% 
  left_join(., protVar, 'Uniprot') %>%
  dplyr::rename(protA = Uniprot, Uniprot = protB, cvA = cv) %>%
  left_join(., protVar, 'Uniprot') %>% dplyr::rename(protB = Uniprot, cvB = cv) %>%
  mutate(totVar = sqrt(cvA^2 + cvB^2))

ggplot(corVect, aes(x = cor, y = totVar)) + geom_hex(bins = 250) + theme_bw() +
  xlim(-1,1) + ylim(0,0.4) + 
  geom_vline(xintercept = 0.7, colour = 'red', linetype = 'dashed') +
  geom_hline(yintercept = 0.2, colour = 'red', linetype = 'dashed')

ggplot(filter(corVect, totVar > 0.2, cor > 0.7), aes(x = cor, y = totVar))  + 
  geom_point(alpha = 0.15, size = 0.1) + theme_bw() + xlim(-1,1) + ylim(0,0.4)
```

## Total Variance
```{r}
protImputeTidy <- protImpute %>% 
  gather(key = 'cell', value = 'abundance', 2:ncol(protImpute)) %>%
  separate(cell, into = c('run','label','condition'), remove = F) %>% 
  filter(condition != 'ctrl')

### Prototype function
## Combine the apply find correlation value and pulling out the CV value
geneList <- allGO$Uniprot[9] %>% unlist()
protCombs <- t(combn(geneList, 2))

groupVar <- function(x){
  geneList <- unlist(x)
  groupVars <- protVar %>% filter(Uniprot %in% geneList)
  return(sqrt(sum(groupVars$cv^2)))
}

# Fix by including new code from totVarGroup()
groupCor <- function(x){
  geneList <- unlist(x)

  if (length(geneList) > 1) {
    protCombs <- t(combn(geneList, 2))
    return(sqrt(sum(totVarPair(protCombs))))
  } else{
    return(as.numeric(protVar[protVar$Uniprot == geneList[1], 2]))
  }
}

totVarPair <- function(x){
  A <- as.character(x[1])
  B <- as.character(x[2])
  varA <- as.numeric(protVar[protVar$Uniprot == A, 2])
  varB <- as.numeric(protVar[protVar$Uniprot == B, 2])
  corAB <- cor[A,B]
  tot <- varA^2 + varB^2 + 2*corAB*varA*varB
  return(tot)
}

totVarGroup <- function(x){
  geneList <- unlist(x)

  if (length(geneList) > 1) {
    protCombs <- t(combn(geneList, 2))
    return(sqrt(sum(apply(protCombs, 1, function(x) totVarPair(x)))))
    #return(sqrt(sum(totVarPair(protCombs))))
  } else{
    return(as.numeric(protVar[protVar$Uniprot == geneList[1], 2]))
  }

}

allGO1 <- allGO %>% filter(!is.na(GOterms)) %>% rowwise() %>% 
  mutate(n = length(unlist(Uniprot)),
         totVar = totVarGroup(Uniprot),
         normVar = totVar / n) %>%
  filter(n > 3)

allGO1 %>% arrange(desc(normVar)) %>% print(n = 100)
allGO1 %>% arrange(normVar) %>% print(n = 100)

allGO1$GO <-factor(allGO1$GO, levels = allGO1$GO[order(allGO1$normVar)])

ggplot(allGO1, aes(x = normVar, y = totVar)) + geom_point()
ggplot(allGO1, aes(x = log2(n), y = log2(totVar))) + geom_point()
ggplot(allGO1, aes(x = GO, y = normVar)) + geom_point()

ggplot(allGO1, aes(x = normVar)) + geom_histogram(bins = 30) + 
  geom_vline(xintercept = c(0.15, 0.1), colour = 'red', linetype = 'dashed') +
  theme_bw()

```

### Abundance based on intensity
Does expression level explain lower variance?
```{r}

ggplot(protRawMean, aes(x = log2(mean), y = noise)) + geom_point() + theme_bw()

```

It doesn't appear to be the case. This could be because these are likely the top ~1000 highest expressed proteins that have a dynamic range of ~250X.

### Move up the GO branch to see how many high/low variance GO terms are from which parent branch
```{r}

```

### Is lower variance compensating for high correlations? How much of a role is negative correlation playing in reducing the total variance?
```{r}
dat <- left_join(ungroup(test), allGO1, by = "GOterms") %>%
  dplyr::select(GOterms, GO.x, ontology.x, Uniprot.x, cors.x, cv.x, n.x,medCors, medCV, normVar)

dat$score <- 'Med'
dat[dat$normVar > 0.145, 'score'] <- 'High'
dat[dat$normVar < 0.115, 'score'] <- 'Low'

ggplot(dat, aes(x = medCors, y = medCV, colour = score)) + geom_point() 
ggplot(dat, aes(x = medCors, y = medCV)) + geom_point() + geom_smooth(method = 'lm') + facet_grid(~score)

ggplot(filter(dat, n.x > 5), aes(x = medCors, y = medCV, colour = score)) + 
  geom_point(alpha = 0.75) + scale_color_manual(values=c("blue", "red", "grey"))
# + #scale_colour_gradient(low="red", high="blue") +
#  geom_text_repel(data = filter(dat, score != 'Med'), aes(label = GOterms), colour = 'black', size = 1)

ggplot(filter(dat, n.x > 5, score != 'Med', ontology.x != 'CC'), aes(x = medCors, y = medCV, colour = normVar)) + 
  geom_point() + scale_colour_gradient(low="red", high="blue") + 
  geom_smooth(method = 'lm') + facet_wrap(~score, scales = 'free') +
  geom_text_repel(data = filter(dat, n.x > 5, score != 'Med'), aes(label = GO.x), colour = 'black', size = 2)

plot_ly(filter(dat, n.x > 5), z = ~normVar, y = ~medCV, x = ~medCors, color = ~score, colors = c("red", "blue", "grey")) %>%
  add_markers() %>%
  layout(scene = list(zaxis = list(title = 'Normalized Variance'),
                     yaxis = list(title = 'Median CV'),
                     xaxis = list(title = 'Median Correlation')))

ggplot(filter(dat, n.x > 5), aes(y = normVar, x = medCV)) + geom_point(alpha = 0.75)
ggplot(filter(dat, n.x > 5), aes(y = normVar, x = medCors)) + geom_point(alpha = 0.75)

library(GGally)
p <- ggpairs(filter(dat, n.x > 5), columns = c('medCV', 'medCors','normVar'), 
        ggplot2::aes(colour = score, alpha = 0.5))

for(i in 1:p$nrow) {
  for(j in 1:p$ncol){
    p[i,j] <- p[i,j] + 
        scale_fill_manual(values=c("blue", "red", "grey")) +
        scale_color_manual(values=c("blue", "red", "grey"))  
  }
}
p

```

## Variance within highly correlated GO terms
```{r message=FALSE, warning=FALSE}
allGO[grep('large ribo',allGO$GO),]
hCor <- filter(allGO, GOterms == 'GO:0015934')$Uniprot %>% unlist()

hCor <- protImpute %>% filter(Uniprot %in% hCor) %>% 
  mutate(symbol = mapIds(org.Xl.eg.db, keys = Uniprot, keytype = 'UNIPROT', column = 'SYMBOL')) %>%
  dplyr::select(Uniprot, symbol, everything())

hCor <- hCor %>% gather(key = sample, value = abundance, 3:ncol(hCor)) %>%
  dplyr::select(sample, symbol, abundance) %>% spread(key = symbol, value = abundance) %>%
  gather(key = protein, value = abundance, 2:ncol(.))

hCorSD <- hCor %>% group_by(sample) %>% summarise(sd = sd(abundance))

ggplot(hCorSD, aes(x = sample, y = sd)) + geom_point() + geom_hline(yintercept = sd(hCor$abundance))

BC_delta <- function(x) {
  hCor <- filter(allGO, GOterms == x)$Uniprot %>% unlist()

  hCor <- protImpute %>% filter(Uniprot %in% hCor) %>%
    mutate(symbol = mapIds(
      org.Xl.eg.db,
      keys = Uniprot,
      keytype = 'UNIPROT',
      column = 'SYMBOL'
    )) %>%
    dplyr::select(Uniprot, symbol, everything())

  hCor <- hCor %>% gather(key = sample, value = abundance, 3:ncol(hCor)) %>%
    dplyr::select(sample, symbol, abundance)

  hCorSD <-
    hCor %>% group_by(sample) %>% summarise(sd = sd(abundance))

  return(hCorSD %>% summarise(sum(sd < sd(hCor$abundance)) / n()) %>% as.numeric())
}

dat <- dat %>% rowwise() %>% mutate(bcDelta = BC_delta(GOterms))

ggplot(dat, aes(x = medCors, y = bcDelta)) + geom_point(alpha = 0.5) + 
  geom_text_repel(data = filter(test, medCors > 0.2), 
                    aes(label = GO), colour = 'black', size = 2)
```