### 2020-03-02 Untitled.Rmd
improved analysis

---
title: "Single Cell Xenopus TMT Analysis"
author: "Kyle Kovary"
date: "5/15/2019"
output: pdf_document
---

## Setup
```{r}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(RColorBrewer)
library(pheatmap)
library(GGally)
library(VIM)
library(org.Xl.eg.db)
library(GO.db)
library(ggrepel)
library(plotly)
```

## Import Data
```{r setup, include=FALSE}

# Import Data
files = list.files('TMTdata/', pattern = 'TargetPeptideSpectrumMatch', full.names = T)

for(i in seq_along(files)){
  if(i == 1 & exists('pepRaw')){
    rm(pepRaw)
  } 
  if(!exists('pepRaw')){
    pepRaw = read_tsv(file = files[i]) %>% mutate(file = files[i], massID = 1:nrow(.))
  } else{
    temp = read_tsv(file = files[i]) %>% mutate(file = files[i], massID = 1:nrow(.))
    colnames(temp) = colnames(pepRaw)
    pepRaw = rbind(pepRaw, temp)
    rm(temp)
  }
}
pepRaw <- pepRaw %>% dplyr::select(file, massID, everything()) %>% filter(!is.na(`126`), 
                                                                  `Quan Usage` == 'Use',
                                                                  Confidence != 'Low')

```

## Data Organization
```{r }
pepTidy <- pepRaw %>% 
  mutate(run = c('ctrl','a','b','c','d','e')[match(pepRaw$file, files)]) %>% 
  group_by(file, `Annotated Sequence`) %>% 
  mutate(pepID = letters[1:n()]) %>% ungroup() %>%
  dplyr::select(file, run, pepID, everything()) %>%
  gather(key = label, value = intensity, '126':'131') %>% 
  dplyr::select(run, label, intensity, everything(), -file) %>% 
  mutate(condition = ifelse(run == 'ctrl', 'ctrl',c('ctrl',0,20,40,60,80)[match(label,c(126:131))]),
         intensity = as.numeric(intensity)) %>% unite(peptide, c(`Annotated Sequence`,'pepID')) %>%
  dplyr::rename(Uniprot = `Master Protein Accessions`) %>% 
  unite(cell, c('run','label','condition'), remove = F) %>%
  dplyr::select(cell, run, label, condition, peptide, Uniprot, intensity)
```

## Normalization
```{r}
protImpute <- pepTidy %>% 
  group_by(peptide,run) %>% 
  mutate(intensity = intensity / intensity[label == "126"]) %>%
  ungroup() %>%
  filter(run != "ctrl", condition != "ctrl") %>%
  #dplyr::filter(abs((intensity)) < 2*sd((pepTidy$intensity), na.rm = T)) %>%
  group_by(cell, run, label, condition, Uniprot) %>% 
  summarise(abundance = mean(intensity, na.rm = T)) %>% ungroup() %>%
  group_by(Uniprot) %>%
  mutate(abundance = abundance / mean(abundance, na.rm = T)) %>% 
  group_by(cell) %>% 
  mutate(abundance = abundance / mean(abundance, na.rm = T)) %>% 
  ungroup() %>% 
  dplyr::select(cell, Uniprot, abundance) %>% 
  spread(cell, abundance) %>%
  kNN(data = ., k = 1, impNA = TRUE) %>% as_tibble() %>% dplyr::select(Uniprot:e_131_80)

```

## Summary Stats
```{r}
protVar <- protImpute %>% gather(key = 'cell', value = 'abundance', 2:ncol(protImpute)) %>%
  separate(cell, into = c('run','label','condition'), remove = F) %>% 
  filter(condition != 'ctrl') %>% group_by(Uniprot) %>%
  summarise(var = var(abundance),
            sd = sd(abundance),
            mean = mean(abundance),
            cv = sd / mean)

cor = cor(t(log2(protImpute[c(1:ncol(protImpute))[-c(1,grep('ctrl',colnames(protImpute)))]])), use = 'pairwise.complete.obs')
colnames(cor) = protImpute$Uniprot
rownames(cor) = protImpute$Uniprot

cov = cov(t(log2(protImpute[c(1:ncol(protImpute))[-c(1,grep('ctrl',colnames(protImpute)))]])), use = 'pairwise.complete.obs')
colnames(cov) = protImpute$Uniprot
rownames(cov) = protImpute$Uniprot
```

## Map GO Terms to Proteins

```{r eval=TRUE, message=FALSE, warning=FALSE}
corVector <- function(uniprotList, corMat){
  targets = uniprotList %>% unlist()
  temp = corMat[rownames(corMat) %in% targets,colnames(corMat) %in% targets]
  temp[lower.tri(temp, diag = F)]
}

selectCV <- function(x){
  protVar %>% filter(Uniprot %in% unlist(x)) %>% 
    dplyr::select(cv) %>% as_tibble() %>% as.list()
}

keys <- keys(org.Xl.eg.db, keytype = 'UNIPROT')

allGO <- protImpute %>% dplyr::select(Uniprot) %>% filter(Uniprot %in% keys) %>% 
  mutate(GOterms = mapIds(org.Xl.eg.db, 
                     keys = Uniprot, 
                     keytype = 'UNIPROT', 
                     column = 'GOALL', 
                     multiVals = 'list')) %>%
  unnest(cols = GOterms) %>% nest(Uniprot = Uniprot) %>%
  rowwise() %>% mutate(n = length(unlist(Uniprot))) %>% ungroup() %>%
  mutate(GO = mapIds(GO.db, 
                     keys = GOterms, 
                     keytype = 'GOID', 
                     column = 'TERM') %>% as.character(),
         ontology = mapIds(GO.db, 
                     keys = GOterms, 
                     keytype = 'GOID', 
                     column = 'ONTOLOGY') %>% as.character()) %>%
  dplyr::select(GOterms, GO, ontology, everything())%>% 
  group_by(GOterms) %>% 
  dplyr::mutate(cors = corVector(Uniprot, cor) %>% as_tibble() %>% as.list(),
                covs = corVector(Uniprot, cov) %>% as_tibble() %>% as.list(),
                cv = selectCV(Uniprot)) %>% ungroup() %>%
  rowwise() %>% mutate(medCors = mean(cors, na.rm = TRUE),
                       medCV = mean(cv, na.rm = TRUE),
                       totVar = sqrt(sum(cv) + 2*sum(covs)),
                       normVar = totVar / n) %>%
  filter(n > 3) %>%
  ungroup()

```

## Plots
```{r fig.height=8, fig.width=5}
allGO1 <- allGO1 %>% mutate(interval = cut_number(normVar, n = 10))

ggplot(allGO1, aes(x = normVar)) + geom_histogram(bins = 50)

ggplot(allGO1, 
       aes(x = medCors, y = medCV, color = interval == "[0.00652,0.0333]")) + 
  geom_point(alpha = 0.5) + theme_bw() +
  #facet_wrap(~interval) +
  theme(legend.position = "bottom")

ggplot(allGO1, 
       aes(x = medCors, y = medCV)) + 
  geom_point(alpha = 0.5) + theme_bw() +
  geom_label_repel(data = filter(allGO1, 
                                 interval == "[0.00652,0.0333]",
                                 medCors > 0.15), 
                  aes(label = GO), 
                  colour = 'black', 
                  size = 2,
                  force = 10) + 
  theme_bw()
```

## Cor histogram
```{r}
goTerms <- c("GO:0006739","GO:0006096","GO:0015934")

allGO[grep('large ribosomal subunit',allGO$GO),]
glyc <- filter(allGO, GOterms == 'GO:0006096')$Uniprot %>% unlist()

cor_rug_plot <- function(x){
  temp <- filter(allGO, GOterms %in% x) %>%
    dplyr::select(GO, cors) %>% unnest(cols = cors)

  temp <- allGO %>% select(GO, cors) %>%
    unnest(cols = cors) %>% mutate(GO = "all") %>%
    rbind(temp,.)

  ggplot(temp, aes(x = cors, fill = GO, color = GO)) +
    geom_density(alpha = 0.25) + 
    geom_rug(data = filter(temp, GO != "all")) + 
    theme_bw()
}

```

## KEGG
```{r}
library(KEGG.db)
kegg <- plyr::ldply(as.list(org.Xl.egPATH2EG),data.frame) %>% as_tibble() %>%
  rename(kegg_id = ".id", gene_id = "X..i..")

kegg_names <- plyr::ldply(as.list(KEGGPATHNAME2ID),data.frame) %>%
  rename(kegg_name = ".id",
         kegg_id = "X..i..")

kegg <- protImpute %>% dplyr::select(Uniprot) %>% filter(Uniprot %in% keys) %>% 
  mutate(KEGG = mapIds(org.Xl.eg.db, 
                     keys = Uniprot, 
                     keytype = 'UNIPROT', 
                     column = "PATH", 
                     multiVals = 'list'))

allKEGG <- unnest(kegg, cols = KEGG) %>% nest(Uniprot = Uniprot) %>%
  rowwise() %>% mutate(n = length(unlist(Uniprot))) %>% ungroup() %>%
  mutate(KEGG_name = kegg_names$kegg_name[match(KEGG,kegg_names$kegg_id)])

corVector <- function(uniprotList, corMat){
  targets = uniprotList %>% unlist()
  temp = corMat[rownames(corMat) %in% targets,colnames(corMat) %in% targets]
  temp[lower.tri(temp, diag = F)]
}

selectCV <- function(x){
  protVar %>% filter(Uniprot %in% unlist(x)) %>% 
    dplyr::select(cv) %>% as_tibble() %>% as.list()
}

allKEGG <- allKEGG %>% group_by(KEGG) %>% 
  dplyr::mutate(cors = corVector(Uniprot, cor) %>% as_tibble() %>% as.list(),
                covs = corVector(Uniprot, cov) %>% as_tibble() %>% as.list(),
                cv = selectCV(Uniprot)) %>% ungroup() %>%
  rowwise() %>% mutate(medCors = mean(cors, na.rm = TRUE),
                       medCV = mean(cv, na.rm = TRUE)) %>%
  ungroup() %>%
filter(n > 3,
       !is.na(medCors),
       !is.na(medCV))

allKEGG <- allKEGG %>% 
  rowwise() %>% mutate(totVar = sqrt(sum(cv, na.rm = T) + sum(covs, na.rm = T)),
                             normVar = totVar / n) %>% ungroup() %>%
  mutate(interval = cut_number(normVar, n = 4))

ggplot(allKEGG, aes(x = medCors, y = medCV)) + geom_point(aes(color = interval)) +
  geom_label_repel(data = allKEGG, 
                  aes(label = KEGG_name), 
                  colour = 'black', 
                  size = 2,
                  force = 10)

glyc <- filter(allKEGG, KEGG == '00010')$Uniprot %>% unlist()

glycMat <- cor[rownames(cor) %in% glyc,colnames(cor) %in% glyc]
rownames(glycMat) <- mapIds(org.Xl.eg.db, keys = rownames(glycMat), keytype = 'UNIPROT', column = 'SYMBOL')
pheatmap(glycMat, show_colnames = T, show_rownames = T, color = mypal, breaks = seq(-1,1, by = 0.01),
         cutree_cols = 3, cutree_rows = 3)

# pathview(genes, 
#          pathway.id = pathway, 
#          species = 'mmu', 
#          same.layer = F,
#          low = list(gene = 'blue'),
#          mid = list(gene = 'transparent'),
#          high = list(gene = 'red'),
#          limit = list(gene = c(-2,2))
# )

```