### 2020-03-03 Protein Analysis.Rmd

---
title: "Single Cell Xenopus TMT Analysis2"
author: "Kyle Kovary"
date: "5/15/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(RColorBrewer)
library(pheatmap)
library(GGally)
library(VIM)
library(org.Xl.eg.db)
library(GO.db)
library(ggrepel)
library(plotly)
library(ComplexHeatmap)

# Import Data
files = list.files('TMTdata/', pattern = 'TargetProtein', full.names = T)

for(i in seq_along(files)){
  if(i == 1 & exists('protRaw')){
    rm(protRaw)
  } 
  if(!exists('protRaw')){
    protRaw = read_tsv(file = files[i]) %>% mutate(file = files[i], massID = 1:nrow(.))
  } else{
    temp = read_tsv(file = files[i]) %>% mutate(file = files[i], massID = 1:nrow(.))
    colnames(temp) = colnames(protRaw)
    protRaw = rbind(protRaw, temp)
    rm(temp)
  }
}
protRaw <- protRaw %>% dplyr::select(file, massID, everything()) %>% filter(`Confidence A2` != 'Red',
                                                                            Master == 'IsMasterProtein')

```

## Data Organization
```{r }
protRaw = protRaw %>% 
  mutate(run = c('ctrl','a','b','c','d','e')[match(protRaw$file, files)]) %>% 
  dplyr::select(file, run, everything())

protTidy = protRaw %>% 
  gather(key = label, 
         value = intensity, 
         `Ratios: (127) / (126)`:`Ratios: (131) / (126)`) %>% 
  dplyr::select(run, label, intensity, everything(), -file) %>% 
  mutate(condition = c(0,20,40,60,80)[match(label,names(protRaw)[16:20])],
         intensity = as.numeric(intensity)) %>% 
  dplyr::rename(Uniprot = Accession) %>% 
  unite(cell, c('run','label','condition'), remove = F) %>%
  dplyr::select(cell, run, label, condition, Uniprot, intensity)

protTidy <- protTidy[-grep("_",protTidy$Uniprot),]
```

## Data Normalization
```{r }
protTidy <- protTidy %>% group_by(Uniprot) %>% mutate(intensity = intensity / median(intensity, na.rm = TRUE)) %>%
  group_by(cell) %>% mutate(intensity = intensity / median(intensity, na.rm = TRUE)) %>% ungroup()
```

## Impute data
```{r}
protImpute <- protTidy %>% dplyr::select(run,condition,Uniprot,intensity) %>%
  unite(col = "cell", run:condition) %>%
  pivot_wider(names_from = cell, values_from = intensity) %>%
  kNN(data = ., k = 1) %>% as_tibble() %>% dplyr::select(Uniprot:e_80)

protVar <- protImpute %>% gather(key = 'cell', value = 'abundance', 2:ncol(protImpute)) %>%
  separate(cell, into = c('run','label'), remove = F) %>% 
  filter(run != 'ctrl') %>% group_by(Uniprot) %>%
  summarise(var = var(abundance),
            sd = sd(abundance),
            mean = mean(abundance),
            cv = sd / mean)

ggplot(protVar, aes(x = sd)) + geom_histogram(bins = 50) + theme_bw()
ggplot(protVar, aes(x = mean, y = cv)) + geom_point()
```

## OPTIONAL - Filter out proteins with variance less than control
```{r}
bio_vs_tech_var <- protImpute %>% gather(key = 'cell', value = 'abundance', 2:ncol(protImpute)) %>%
  separate(cell, into = c('run','label'), remove = F) %>% 
  mutate(bio = ifelse(run == "ctrl",
                      "tech_var",
                      "bio_var")) %>%
  group_by(Uniprot, bio) %>%
  summarise(cv = sd(abundance) / mean(abundance)) %>%
  pivot_wider(names_from = bio, values_from = cv) %>%
  mutate(keep = bio_var > tech_var)

ggplot(bio_vs_tech_var, aes(x = log2(bio_var / tech_var))) + geom_histogram(bins = 50) +
  geom_vline(xintercept = 0)

keep <- bio_vs_tech_var %>% filter(keep) %>% pull(Uniprot)
```

## Variance vs Mean
```{r}
ggplot(protVar, aes(x = log2(mean), y = var/mean^2)) + geom_point() + geom_smooth(method = lm, formula = y ~ poly(x, 2))
```

## Calculate variance and covariance
```{r}
cor = cor(t(log2(protImpute[c(1:ncol(protImpute))[-c(1,grep('ctrl',colnames(protImpute)))]])), use = 'pairwise.complete.obs')
colnames(cor) = protImpute$Uniprot
rownames(cor) = protImpute$Uniprot

cov = cov(t(log2(protImpute[c(1:ncol(protImpute))[-c(1,grep('ctrl',colnames(protImpute)))]])), use = 'pairwise.complete.obs')
colnames(cov) = protImpute$Uniprot
rownames(cov) = protImpute$Uniprot

mypal = rev(colorRampPalette(brewer.pal(11, 'RdBu'))(201))
k = 2
# map = pheatmap(cor, show_colnames = F, show_rownames = F, color = mypal, breaks = seq(-1,1, by = 0.01), cutree_cols = k, cutree_rows = k, legend = F)
#map

tibble(Uniprot = protImpute$Uniprot, Cluster = cutree(map$tree_row, k = k)) %>% write_csv('/Users/kylekovary/Downloads/Map.csv')
```

## PCA Analysis
```{r}
remove = which(is.na(pairs), arr.ind = T)[,1] %>% unique()
pca = prcomp(protImpute[,2:ncol(protImpute)])

pcaDF = tibble(cell = rownames(pca$rotation), PC1 = pca$rotation[,1], PC2 = pca$rotation[,2])
pcaDF = pcaDF %>% separate(cell, into = c('run','label'), remove = F) %>% print()
ggplot(pcaDF, aes(x = PC1, y = PC2, colour = run)) + geom_point(size = 5)
```

## Variance and Co-variance Between Cells After Cell Normalization

```{r echo=TRUE, fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
protTidy = protTidy %>% group_by(cell) %>% mutate(intensity = intensity / median(intensity, na.rm = T))

# ggplot(protTidy, aes(log2(intensity), colour = cell)) + geom_density() + theme_bw() + xlim(-8,8)
```

```{r echo=TRUE, fig.height=10, fig.width=10,message=FALSE, warning=FALSE}
pairs = protTidy %>% filter(condition == 'ctrl') %>% 
  dplyr::select(cell, peptide, intensity) %>% spread(cell, intensity)

# ggpairs(log2(pairs[,2:ncol(pairs)]), lower = list(continuous = wrap("points", alpha = 0.1, size = 0.1))) + theme_bw()
```

## Variance and Co-variance Between Cells After Peptide Batch Normalization

```{r echo=TRUE, fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
protTidy = protTidy %>% group_by(run, peptide) %>% 
  mutate(intensity = intensity / median(intensity, na.rm = T)) %>% ungroup()

# ggplot(protTidy, aes(log2(intensity), colour = cell)) + geom_density() + theme_bw() + xlim(-8,8)
```

```{r echo=TRUE, fig.height=10, fig.width=10,message=FALSE, warning=FALSE}
pairs = protTidy %>% dplyr::filter(condition == 'ctrl') %>% 
  dplyr::select(cell, peptide, intensity) %>% spread(cell, intensity)

# ggpairs(log2(pairs[,2:ncol(pairs)]), lower = list(continuous = wrap("points", alpha = 0.1, size = 0.1))) + theme_bw()
```