### 2020-03-18 Manuscript/test-tikz.tex

% Created by tikzDevice version 0.12.3 on 2020-03-05 17:30:43
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{amsmath}
\usepackage{tikz}
\usepackage{xcolor}
\usetikzlibrary{calc}
\usepackage[active,tightpage,psfixbb]{preview}
\PreviewEnvironment{pgfpicture}\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (505.89,505.89);
\begin{scope}
\path[clip] ( 49.20, 61.20) rectangle (480.69,456.69);
\definecolor{drawColor}{RGB}{0,0,255}

\path[draw=drawColor,line width= 0.8pt,line join=round,line cap=round] ( 65.18,442.04) --
	( 78.50,436.06) --
	( 91.82,430.08) --
	(105.13,424.10) --
	(118.45,418.12) --
	(131.77,412.14) --
	(145.09,406.16) --
	(158.40,400.17) --
	(171.72,394.19) --
	(185.04,388.21) --
	(198.36,382.23) --
	(211.67,376.25) --
	(224.99,370.27) --
	(238.31,364.29) --
	(251.63,358.31) --
	(264.94,352.32) --
	(278.26,346.34) --
	(291.58,340.36) --
	(304.90,334.38) --
	(318.22,328.40) --
	(331.53,322.42) --
	(344.85,316.44) --
	(358.17,310.46) --
	(371.49,304.48) --
	(384.80,298.49) --
	(398.12,292.51) --
	(411.44,286.53) --
	(424.76,280.55) --
	(438.07,274.57) --
	(451.39,268.59) --
	(464.71,262.61);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 51.86, 61.20) -- (451.39, 61.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 51.86, 61.20) -- ( 51.86, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (118.45, 61.20) -- (118.45, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (185.04, 61.20) -- (185.04, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (251.63, 61.20) -- (251.63, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (318.22, 61.20) -- (318.22, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (384.80, 61.20) -- (384.80, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (451.39, 61.20) -- (451.39, 55.20);
\end{scope}
\end{tikzpicture}

\end{document}